---
title: "R - Assignment Mod 2 Week 3 - Pizza EDA"
author: "Ricardo Diaz"
date: "today"
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
library(tidyverse)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r}
# 1. Answer each question using in words/paragraph. 
# 2. DO NOT use comments inside code blocks (like here) to answer anything. Those are for notes between coders/self. They 
# will be ignored, and not counted as answers. 
# 3. Keep the line/option    knitr::opts_chunk$set(warning = F, results = "hide", message = F)
# You can temporarily turn this on/off and use other option while you are working on the homework if it helps. 
# The submitted work should have this option selected instead.
# 4. All charts/graphs/tables should have appropriate captions. 
```

# HW Assignment - EDA

This pizza ingredient dataset is from data.world (@sdhilip) The variables are:  

Variable | Definition  
-----|------- 
brand	 | Pizza brand
id | ID
mois | Amount of water per 100 grams in the sample
prot | Amount of protein per 100 grams in the sample
fat | Amount of fat per 100 grams in the sample
ash | Amount of ash per 100 grams in the sample
sodium | Amount of sodium per 100 grams in the sample
carb | Amount of carbohydrates per 100 grams in the sample
cal | Amount of calories per 100 grams in the sample

As with all your work in this class, knit the RMD file into HTML, zip it with the RMD, and submit the zip file on Blackboard. 

Compose your answers using inline R code instead of using the code-block output as much as you can. Coder's comments inside code blocks are never graded. 

## Pizza Ingredient dataset

### Question 1  
**Import the dataset into R.**   
Data import to R

```{r}
pizza = data.frame(read.csv("Pizza.csv"))
```

### Question 2  
**How many data points are there?**  
There are 300 data points

```{r}
nrow(pizza)
```



### Question 3  
**Look at the summary statistics of the dataset.**  
Summary statistics of the pizza dataset

```{r}
xkablesummary(pizza)
```


### Question 4 
**Make Plots**  
For sodium and calorie, make histograms, boxplots (using `ggplot()`), and QQ-plots (just regular `qqnorm()` function). Make sure all plots have appropriate titles, x- and y- labels, units on the axes if applicable. It is also much nicer to add some color to your charts instead of plain old black and white. For chart titles, if no appropriate title you can think of, just use y vs x. Don’t get mixed up (somehow I find up to half of the presentations have the wrong ones). It is NEVER x vs y. **Always y vs x**.


```{r}

hist(pizza$sodium, main = "Histogram of Sodium", col.main = "blue", xlab = "Sodium")
hist(pizza$cal, main = "Histogram of Calorie", col.main = "red", xlab = 'Calories')
ggplot(pizza,aes(sodium)) + geom_boxplot( fill = 'blue') + labs(title = 'Sodium Boxplot')
ggplot(pizza,aes(cal)) + geom_boxplot(fill = 'blue') + labs(title = 'Calories Boxplot')
qqnorm(pizza$sodium, main = "Sodium Q-Q Plot", ylab="Sodium Quantiles", col ='green') 
qqnorm(pizza$cal, main = "Calories Q-Q Plot", ylab="Calories Quantiles", col = 'green') 
```


##ggplot(pizza,aes(as.factor(sodium),cal)) + geom_boxplot() 
##ggplot(pizza,aes(sodium,cal)) + geom_boxplot(fill = 'light blue') + labs(tile = 'Calories vs Sodium Boxplot')
##+ labs( x = "N of Sodium, )
##hist(new_baseball$weight,main = "Normal Distribution Players Weight ", col.main = "green")
##str(pizza)
##ggplot(new_baseball,aes(x = weight,y = height,colour = Team))+
##geom_point(size = 2) +
##labs(title = "Plot Graph Weight vs Players")


### Question 5 
**Outliers**  
Use the `ezids::outlierKD2()` function to remove the outliers for sodium, then run the function again to remove outliers on calories. Re-do the QQ-plots for these two variables. Do you see much improvements?

```{r}
outlierKD2(pizza,sodium)
outlierKD2(pizza,cal)
qqnorm(pizza$sodium, main = "Sodium Q-Q Plot", ylab="Sodium Quantiles", col ='green') 
qqnorm(pizza$cal, main = "Calories Q-Q Plot", ylab="Calories Quantiles", col = 'green') 
```
There is an improvement because there is no gap butit makes Q-Q plots harder to understand

