---
title: "Mini Project Team 2"
author: "Team 2"
date: "today"
# date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r init, include=FALSE}
# some of common options (and the defaults) are: 
# include=T, eval=T, echo=T, results='hide'/'asis'/'markup',..., collapse=F, warning=T, message=T, error=T, cache=T, fig.width=6, fig.height=4, fig.dim=c(6,4) #inches, fig.align='left'/'center','right', 
library(ezids)
# knitr::opts_chunk$set(warning = F, results = "markup", message = F)
knitr::opts_chunk$set(warning = F, results = "hide", message = F)
options(scientific=T, digits = 3) 
# options(scipen=9, digits = 3) 
# ‘scipen’: integer. A penalty to be applied when deciding to print numeric values in fixed or exponential notation.  Positive values bias towards fixed and negative towards scientific notation: fixed notation will be preferred unless it is more than ‘scipen’ digits wider.
# use scipen=999 to prevent scientific notation at all times
```

```{r loanpredict}
getwd()
loanpredict <- read.csv("Training Data.csv", header = TRUE)
str(loanpredict)
xkabledplyhead(loanpredict)
```
## Exploratory Data Analysis 

1. Basic statistics
    + mean, s.d., median, range (four spaces at the start of line for proper sub-list indentation)
1. Simple correlations and tests
    + correlation matrix if applicable
    + z-test, t-test, anova test if applicable
    + chi-squared test if applicable
1. Normality
    + QQ-plot
    + boxplot
    + histogram 
    + Shapiro-Wilk test
    + ...
    
```{r EDA_basicstats, result = asis}
xkablesummary(loanpredict)
```
#### Basic Statistics
```{r}
mean(loanpredict$Income)
mean(loanpredict$Age)
```


```{r}
```
#### Correlations and Tests

### Normality check/test

#### Histogram

#### QQ-plot

#### Boxplot

#### Shapiro-Wilk test

## Linear Model

### Findings

### Plot 3D

# Conclusion


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
